### 2017-11-13 Draft2/Appending.Growth.tex
Work

Let $\FuncWritesByGrow(n)$ denote the total number of writes made by $\FuncGrow$, and let $\VarNumItemsTail$ be the size of the tail. Since Corollary \ref{coro:GrowthSeq} also holds true for growth arrays, $\FuncGrow$ is called $|\VarGrowSeq|$ times. A buffer is appended to the tail each time $\FuncGrow$ is called. Thus, the tail's size is
\begin{align*}
\VarNumItemsTail &= |\VarGrowSeq| = |\VarCapacitySeq| - 1 = \VarUseful
\end{align*}
Then the formula for $\FuncWritesByGrow(n)$ is simply $\FuncWritesTail(\VarUseful)$, where $\FuncWritesTail$ denotes the tail's write cost function (that is, the write cost function for dynamic arrays). Finally, adding the writes made by $\FuncAppend$, the formula for $\FuncWrites(n)$ is
\begin{align*}
\FuncWrites(n) = n + \FuncWritesTail(\VarUseful)
\end{align*}
Now, I approximate $\FuncWrites(n)$ using $\sim$. To do this, I will derive the big-O complexity of $\VarUseful$.

@@

\begin{proof}
	From Lemma \ref{lem:CapacitySeq}, $\VarUseful = \max(\ExprMessy, 0)$. As mentioned in Lemma \ref{lem:ToVarUsefulPowerBounds}, for sufficiently large $n$, $\VarUseful = \ExprMessy$. Then
	\begin{align*}
	\VarUseful \sim \ExprMessy \sim \left( \ExprMessyInner \right) \sim \log_{\VarGrowthFactor} n
	\end{align*}
	By [lemma], $O(\VarUseful) = O(\log_{\VarGrowthFactor} n) = O(\log n)$ as desired.
\end{proof}

Now when $\FuncWrites(n)$ is approximated with $\sim$, the $\FuncWritesTail(\VarUseful)$ term disappears:
\begin{align*}
\FuncWrites(n) &= n + \FuncWritesTail(\VarUseful) = n + O(\VarUseful) = n + O(\log n) \sim n
\end{align*}
The last step is true because of Theorem \ref{thm:RemovesLowerOrderTerms}.

@@

Unlike dynamic arrays, growth arrays never throw away buffers. This means that if the current capacity is $c$, then the total length of buffers allocated to store items is also $c$. Typically, however, $\FuncSpace(n) > c$. This is because growth arrays not only store items in buffers, they also store \textbf{references} (or \textbf{pointers}) to the buffers holding the items, in the tail. Thus, the space the tail allocates must also be considered.

We established in [lemma] that $\VarNumItemsTail = \VarUseful$. Then since $\VarCapacitySeq_\VarUseful$ is the space needed to hold items, and $\FuncSpaceTail(\VarUseful)$ is the space the tail needs to hold references, the formula for $\FuncSpace(n)$ is
\begin{align*}
\FuncSpace(n) = \VarCapacitySeq_\VarUseful + \FuncSpaceTail(\VarUseful)
\end{align*}
I now wish to approximate this using $\sim$. First, note that since $\VarCapacitySeq_\VarUseful = \VarGrowthFactor^\VarUseful\VarInitCapacity$, it follows from Lemma \ref{lem:ToVarUsefulPowerBounds} that $n \FluteLeq \VarCapacitySeq_\VarUseful \FluteLess \VarGrowthFactor{n}$. Since $n$ and $\VarGrowthFactor{n}$ are both $O(n)$, [the squeeze theorem] implies that $\VarCapacitySeq_\VarUseful = O(n)$.

>>>>

Let $\FuncWritesByGrow(n)$ denote the total number of writes made by $\FuncGrow$, and let $\VarNumItemsTail$ be the size of the tail. Since Corollary \ref{coro:GrowthSeq} also holds true for growth arrays, $\FuncGrow$ is called $|\VarGrowSeq|$ times. A buffer is appended to the tail each time $\FuncGrow$ is called. Thus, the tail's size is
\begin{align*}
\VarNumItemsTail &= |\VarGrowSeq| = |\VarCapacitySeq| - 1 = \VarUseful
\end{align*}
Then the formula for $\FuncWritesByGrow(n)$ is simply $\FuncWritesTail(\VarUseful)$, since $\VarUseful$ pointers are appended to the tail, which is a dynamic array. Finally, adding the writes made by $\FuncAppend$, the formula for $\FuncWrites(n)$ is
\begin{align*}
\FuncWrites(n) = n + \FuncWritesTail(\VarUseful)
\end{align*}
Now, I approximate $\FuncWrites(n)$ using $\sim$. To do this, I will derive the big-O complexity of $\VarUseful$.

@@

\begin{proof}
	From Lemma \ref{lem:CapacitySeq}, $\VarUseful = \max(\ExprMessy, 0)$. For sufficiently large $n$, $\VarUseful = \ExprMessy$. Using $\sim$ to analyze $\VarUseful$, I receive
	\begin{align*}
	\VarUseful \sim \ExprMessy \sim \left( \ExprMessyInner \right) \sim \log_{\VarGrowthFactor} n
	\end{align*}
	(It is trivial to show that $\left\lceil f(n) \right\rceil \sim f(n)$ if $f$ is unbounded, since $\left\lceil f(n) \right\rceil - f(n)$ is bounded by $1$.)

	By Theorem \ref{thm:SameBigOClass}, $O(\VarUseful) = O(\log_{\VarGrowthFactor} n) = O(\log n)$ as desired.
\end{proof}

Now when $\FuncWrites(n)$ is approximated with $\sim$, $\FuncWritesTail(\VarUseful)$ disappears:
\begin{align*}
\FuncWrites(n) &= n + \FuncWritesTail(\VarUseful) = n + O(\VarUseful) = n + O(\log n) \sim n
\end{align*}
The last step is justified by Theorem \ref{thm:RemovesLowerOrderTerms}.

@@

Unlike dynamic arrays, growth arrays do not throw away buffers. This means that if the current capacity is $c$, then the total length of buffers allocated to store items is also $c$. Typically, however, $\FuncSpace(n) > c$. This is because the tail of growth arrays (which is a dynamic array) also allocates buffers to store buffer pointers. The space the tail allocates must be considered in the formula for $\FuncSpace(n)$.

In the previous section, I established that $\VarNumItemsTail = \VarUseful$. Since $\VarCapacitySeq_\VarUseful$ is the capacity needed to hold $n$ items, and $\FuncSpaceTail(\VarUseful)$ is the space the tail needs to store $\VarUseful$ pointers, I conclude that the formula for $\FuncSpace(n)$ is
\begin{align*}
\FuncSpace(n) = \VarCapacitySeq_\VarUseful + \FuncSpaceTail(\VarUseful)
\end{align*}
($\FuncSpaceTail$ denotes the space cost function for dynamic arrays.)

% Bookmark
% The following following paragraph doesn't work.

Now, I wish to approximate this using $\sim$. Since $\VarCapacitySeq_\VarUseful = \VarGrowthFactor^\VarUseful\VarInitCapacity$, it follows from Lemma \ref{lem:ToVarUsefulPowerBounds} that $n \FluteLeq \VarCapacitySeq_\VarUseful \FluteLess \VarGrowthFactor{n}$. Since $n$ and $\VarGrowthFactor{n}$ are both $O(n)$, [the squeeze theorem] implies that $\VarCapacitySeq_\VarUseful = O(n)$.
